---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
lgr::get_logger("mlr3")$set_threshold("warn")
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(width = 100, datatable.print.class = FALSE)
```

# mlr3changepoint

<!-- badges: start -->
[![r-cmd-check](https://github.com/sleepysnorlax05/mlr3changepoint/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/sleepysnorlax05/mlr3changepoint/actions/workflows/R-CMD-check.yaml)
[![codecov](https://codecov.io/github/sleepysnorlax05/mlr3changepoint/graph/badge.svg?token=AGF9QGIRKY)](https://codecov.io/github/sleepysnorlax05/mlr3changepoint)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![StackOverflow](https://img.shields.io/badge/stackoverflow-mlr3-orange.svg)](https://stackoverflow.com/questions/tagged/mlr3)
[![Mattermost](https://img.shields.io/badge/chat-mattermost-orange.svg)](https://lmmisld-lmu-stats-slds.srv.mwn.de/mlr_invite/)
<!-- badges: end -->

*mlr3changepoint* extends the [mlr3](https://github.com/mlr-org/mlr3) ecosystem with
**weakly supervised change point detection**. It registers a new `"changepoint"` task
type together with the surrounding infrastructure — tasks, learners, and prediction
objects — so that change point detectors can be trained, resampled, and benchmarked
with the standard mlr3 machinery.

Unlike classical, fully supervised detection, *weak supervision* means the targets are
**labelled regions** of a sequence rather than a label for every observation. Each label
is an interval `(start, end, label)` marking a stretch of the sequence that should contain
a given number of changes (e.g. `"1change"`) or a peak. This is the setup used by
[penaltyLearning](https://cran.r-project.org/package=penaltyLearning) and related
peak-detection work.

> **Note**
> This package is under active development as part of Google Summer of Code 2026.
> The API is experimental and may change.

## Installation

*mlr3changepoint* is not yet on CRAN. Install the development version from GitHub:

```r
# install.packages("remotes")
remotes::install_github("sleepysnorlax05/mlr3changepoint")
```

## Feature Overview

The package adds a `"changepoint"` task type to mlr3's reflections, supporting two
flavours of weak labels, selected via `label_type`:

- `"changepoint"` — regions annotated with the number of changes they contain
  (penaltyLearning-style).
- `"peak"` — regions annotated as peak or background (peak-detection style).

The current building blocks:

| Class | Role | Status |
| :--- | :--- | :--- |
| `TaskCpt` | Weakly supervised change point task — one row per labelled sequence | Available |
| `LearnerCpt` | Abstract base class for change point learners | Base class |
| `PredictionCpt` / `MeasureCpt` | Prediction container and evaluation measures | Planned |

## Task Structure

A `TaskCpt` stores **one row per labelled sequence** and uses two special column roles:

- a **`sequence`** column holding the raw signal (a list column, one numeric vector
  per sequence);
- a **`target`** list column, where each entry is a `data.table(start, end, label)` of
  labelled regions.

`task$truth()` returns the labelled regions and `task$label_type` reports whether the
task carries `"changepoint"` or `"peak"` labels.

## Example

```{r example}
library(mlr3changepoint)
library(data.table)

# one row per sequence: the first has two labelled changes, the second none
set.seed(36)
toy = data.table(
  seq = list(
    c(rnorm(100, 18), rnorm(100, 36), rnorm(100, 9)),
    rnorm(100, 0)
  ),
  label = list(
    data.table(start = c(90, 190), end = c(110, 210), label = c("1change", "1change")),
    data.table(start = 1, end = 100, label = "0changes")
  )
)

task = TaskCpt$new(
  id = "toy", backend = toy,
  target = "label", sequence = "seq", label_type = "changepoint"
)
task

# the weak labels: one set of regions per sequence
task$truth()

# which kind of labels this task carries
task$label_type
```

## Bugs, Questions, Feedback

*mlr3changepoint* is a free and open source software project that encourages participation
and feedback. If you have any issues, questions, suggestions, or feedback, please do not
hesitate to open an issue in the
[issue tracker](https://github.com/sleepysnorlax05/mlr3changepoint/issues). In order to
reproduce issues please use a [reprex](https://reprex.tidyverse.org/).

For general questions about the mlr3 ecosystem, the
[mlr3book](https://mlr3book.mlr-org.com/) and the
[mlr-org community channels](https://stackoverflow.com/questions/tagged/mlr3) are good
starting points.
